Body size patterns

### 2019-07-10 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
took season out of main effects and splines

```{r gams for body size - data prep}
speciesIn <- 'bkt'
depVarIn <- 'observedLength'
kIn <- 4 # # of knots for the speces and river smooths

@@

```{r gam models - find best lm, before fitting detailed gams - only need to run for model selection}
# only including river, season, salmon01 becaseu we assume these have categorical effects

mod0 <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

mod1 <- bam(observedLength ~ riverFactor + factor(season) + factor(salmon01) + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

@@

mod3 <- bam(observedLength ~ riverFactor + factor(season) + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

@@

mod5 <- bam(observedLength ~ factor(season) + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

@@

AIC(mod0,mod1,mod2,mod3,mod4,mod5,mod6) %>% rownames_to_column() %>% arrange(AIC)
summary(mod0)
plot(mod0)
gam.check(mod0)

@@

```{r gam structure given best model from above}

AIC(mod0_YDCcr, mod0_YDCc_r, mod0_YDCcr_Cr_CSr_Tr_Fr_TFr, mod0_YDCcr_Cr_CSr_Ts_Fs_TFs,mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_C_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_CS_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_CSr,mod0_YDCcr_C_CS_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_Tsr_Fsr_TFsr,mod0_YDCcr_CSr_Tsr_Fsr_TFsr, mod0_YDCcr_Tsr_Fsr_TFsr) %>% rownames_to_column() %>% arrange(AIC)

@@

```{r predictions}
# BKT
# system.time(
# mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
#                s(ydayCumul, bs = 'cr') +
#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
#                s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
#                s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
#                s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
#                s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
#                s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
#                s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
#                data = dataIn, 
#                method = "REML")
# )

predGridBKT <- expand.grid(cohortRiverFactor = unique(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr$model$cohortRiverFactor),
                           seasonRiverFactor = unique(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr$model$seasonRiverFactor),
                           ydayCumul = seq(250,1000, by = 30),
                           salmon01 = c(0,1),
                           cohortStrengthScaled = 0, #seq(-1.5,1.5, by = 0.5),
                           meanTemperatureScaledBySeason = 0, #seq(-1.5,1.5, by = 0.5),
                           meanFlowScaledBySeason = 0) %>% #seq(-1.5,1.5, by = 0.5)) %>%
                          # meanTemperatureFlowScaledBySeason = seq(-1.5,1.5, by = 0.5)) %>%
                    mutate(cohortFactor = as.numeric(substr(cohortRiverFactor,1,4)),
                           riverFactor = substring(cohortRiverFactor,6),
                           season = substring(seasonRiverFactor,1),
                           #cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
                           #seasonRiverFactor = paste(season, riverFactor, sep = '_'),
                           #cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortFactor,
                           #river = riverFactor,
                           meanTemperatureFlowScaledBySeason = meanTemperatureScaledBySeason * meanFlowScaledBySeason)

@@

yDayAll <- data.frame(y = 250:1000)
# pick up here
#### need to do something like this  left_join(yDayAll, ydayCumulSeason, by = c('yDay' = 'minY'))

predGridBKT <- expand.grid(riverFactor = unique(dataIn$riverFactor), 
                           cohortFactor = unique(dataIn$cohort),
                           ydayCumul = seq(250,1000, by = 30),
                           salmon01 = c(0,1),
                           cohortStrengthScaled = 0, #seq(-1.5,1.5, by = 0.5),
                           meanTemperatureScaledBySeason = 0, #seq(-1.5,1.5, by = 0.5),
                           meanFlowScaledBySeason = 0) %>%
                    left_join(., ydayCumulSeason) %>%
                    mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
                           seasonRiverFactor = paste(season, riverFactor, sep = '_'),
                           cohort = cohortFactor,
                           #river = riverFactor,
                           meanTemperatureFlowScaledBySeason = meanTemperatureScaledBySeason * meanFlowScaledBySeason)

predBKT <- predict(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr, newdata = predGridBKT)#, type = 'lpmatrix')

ggplot(predGrid2BKT %>% filter(salmon01 == 0),

       aes(ydayCumul, pred, group = cohort, color = cohort)) +
  #geom_point() + 
  geom_vline(xintercept = c(365,365*2), color = 'grey60', linetype = 2) +
  geom_line(size = 1.1) +
  labs(y = 'Predicted body size (mm)',
       x = 'Days since age-0 Jan 1',
       color = 'Cohort') +
  scale_color_viridis_c(breaks = c(2000,2004,2008,2012), labels = c(2000,2004,2008,2012)) +
  theme_publication() +
  theme(legend.position = "right", 
        legend.direction = "vertical",
        legend.title = element_text(face = "plain")) +
  guides(color = guide_colorbar(barheight = 10,
                                barwidth = 1.5,
                                ticks = FALSE,
                                nbin = 100)) +
#  guide_legend(title = 'Cohort') +
  facet_grid(~riverFactor, scales = 'free_y')

@@

#using salmon01 * river because we need diff effects per river - even though AIC is worse for this model
modBKT <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) +
               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn),
             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")

@@

#plot(modBKT, pages = 8, scale = 0)

predGridBKT <- expand.grid(cohortRiverOrdered = unique(modBKT$model$cohortRiverOrdered),
                           ydayCumul = seq(250,1000, by = 30),
                           salmon01 = c(0,1),
                           cohortStrengthScaledBKT = seq(-2,2, by = 0.1)) %>%
                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
                           riverOrdered = substring(cohortRiverOrdered,6),
                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                           cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortOrdered,
                           river = riverOrdered)

@@

modBNT <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = metaPop01) +
               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn) +
               s(cohortStrengthScaledATS, bs = 'cr', k = kIn), 
             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bnt'), method = "REML")

@@

predGridBNT <- expand.grid(cohortRiverOrdered = unique(modBNT$model$cohortRiverOrdered),
                           ydayCumul = seq(250,1000, by = 30),
                           salmon01 = c(0,1),
                           metaPop01 = c(0,1),
                           cohortStrengthScaledBKT = seq(-2,2, by = 0.25),
                           cohortStrengthScaledBNT = seq(-2,2, by = 0.25),
                           cohortStrengthScaledATS = seq(-2,2, by = 0.25)) %>%
                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
                           riverOrdered = substring(cohortRiverOrdered,6),
                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                           cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortOrdered,
                           river = riverOrdered)

@@

predGrid2Both$riverGG <- factor(predGrid2Both$riverOrdered,
                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
                         labels = c("West Brook","Open Large","Open Small","Isolated Small"), ordered = T)
predGrid2Both$speciesGG <- factor(predGrid2Both$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

@@

# cohorts by river/species
ggplot(predGrid2Both %>% 
         filter(
         (cohortStrengthScaledBKT == 0 & salmon01 == 0) | (cohortStrengthScaledBKT == 0 & cohortStrengthScaledBNT == 0 & cohortStrengthScaledATS == 0 & metaPop01 == 1 & salmon01 ==0)),
#           cohortStrengthScaledBKT == 0, cohortStrengthScaledBNT == 0, cohortStrengthScaledATS == 0, metaPop01 == 1, salmon01 == 0),
       aes(ydayCumul, pred, group = cohort, color = cohort)) +
  #geom_point() + 
  geom_vline(xintercept = c(365,365*2), color = 'grey60', linetype = 2) +
  geom_line(size = 1.1) +
  labs(y = 'Predicted body size (mm)',
       x = 'Days since age-0 Jan 1',
       color = 'Cohort') +
  scale_color_viridis_c(breaks = c(2000,2004,2008,2012), labels = c(2000,2004,2008,2012)) +
  theme_publication() +
  theme(legend.position = "right", 
        legend.direction = "vertical",
        legend.title = element_text(face = "plain")) +
  guides(color = guide_colorbar(barheight = 10,
                                barwidth = 1.5,
                                ticks = FALSE,
                                nbin = 100)) +
#  guide_legend(title = 'Cohort') +
  facet_grid(speciesGG ~ riverGG, scales = 'free_y')

@@

>>>>

```{r gams for body size - data prep}
speciesIn <- 'bnt'
depVarIn <- 'observedLength'
kIn <- 4 # # of knots for the speces and river smooths

@@

```{r gam models - find best lm, before fitting detailed gams - only need to run for model selection}
# only including river, salmon01 becaseu we assume these have categorical effects

# not including season
# models with season have big dip in predictions for season 1. Season is...

@@

@@

@@

AIC(mod2,mod4,mod6) %>% rownames_to_column() %>% arrange(AIC)
summary(mod6)
plot(mod6)
gam.check(mod6)

# bkt
#   rowname       df      AIC
# 1    mod6 14.99337 429769.2
# 2    mod2 25.99569 432704.3
# 3    mod4 13.98558 433764.8

#bnt
#   rowname       df      AIC
# 1    mod6 12.99685 238636.2
# 2    mod2 17.99755 242840.1
# 3    mod4 11.98667 243594.3

@@

```{r models with no fixed season effects or categories for splines}

system.time(
mod6_YDCcr_Cr_CSr_Tr_Fr_TFr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod6_YDCcr_Cr_CSr_T_F_TF <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod6_YDCcr_Cr_CS_T_F_TF <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod6_YDCcr_C_CS_T_F_TF <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod6_YDC_C_CS_T_F_TF <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
               data = dataIn, 
               method = "REML")
)

AIC(mod6_YDC_C_CS_T_F_TF,mod6_YDCcr_C_CS_T_F_TF,mod6_YDCcr_Cr_CS_T_F_TF,mod6_YDCcr_Cr_CSr_T_F_TF,mod6_YDCcr_Cr_CSr_Tr_Fr_TFr)  %>% rownames_to_column() %>% arrange(AIC)

summary(mod6_YDCcr_Cr_CSr_Tr_Fr_TFr)
plot(mod6_YDCcr_Cr_CSr_Tr_Fr_TFr, pages = 8, shade = TRUE, scale = 0)

#BKT
#                       rowname        df      AIC
# 1 mod6_YDCcr_Cr_CSr_Tr_Fr_TFr 194.27159 402517.3
# 2    mod6_YDCcr_Cr_CSr_T_F_TF 178.16547 402696.5
# 3     mod6_YDCcr_Cr_CS_T_F_TF 170.29804 403228.7
# 4      mod6_YDCcr_C_CS_T_F_TF 166.30172 404261.2
# 5        mod6_YDC_C_CS_T_F_TF  29.89126 412576.2

#BNT
#                       rowname        df      AIC
# 1 mod6_YDCcr_Cr_CSr_Tr_Fr_TFr 102.82522 225950.4
# 2    mod6_YDCcr_Cr_CSr_T_F_TF  98.12051 225974.0
# 3     mod6_YDCcr_Cr_CS_T_F_TF  96.51127 225994.7
# 4      mod6_YDCcr_C_CS_T_F_TF  92.93128 226010.4
# 5        mod6_YDC_C_CS_T_F_TF  27.97005 231079.5

@@

```{r predictions}

@@

yDayAll <- data.frame(ydayCumul = 250:1000)
yDayAll$season <- NA

# fill in season for each day of the year (yDayAll) based on ydayCumulSeason min values
# need this to get season into all ydayCumul in predGridBKT
for(i in 1:nrow(yDayAll)) {
  for(j in 1:(nrow(ydayCumulSeason) - 1)) {
    if (between(yDayAll$ydayCumul[i], ydayCumulSeason$minY[j], ydayCumulSeason$minY[j + 1])) {
      yDayAll$season[i] <- ydayCumulSeason$season[j]
    }
  }
}

predGrid <- expand.grid(riverFactor = unique(dataIn$riverFactor), 
                           cohortFactor = unique(dataIn$cohort),
                           ydayCumul = seq(250,1000, by = 10),
                           salmon01 = c(0,1),
                           cohortStrengthScaled = 0, #seq(-1.5,1.5, by = 0.5),
                           meanTemperatureScaledBySeason = 0, #seq(-1.5,1.5, by = 0.5),
                           meanFlowScaledBySeason = 0) %>%
                    left_join(., yDayAll) %>%
                    mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
                           seasonRiverFactor = paste(season, riverFactor, sep = '_'),
                           cohort = cohortFactor,
                           #river = riverFactor,
                           meanTemperatureFlowScaledBySeason = meanTemperatureScaledBySeason * meanFlowScaledBySeason)

# predGridBKT2 <- expand.grid(riverFactor = unique(dataIn$riverFactor), 
#                            
#                            ydayCumul = seq(250,1000, by = 10),
#                            salmon01 = c(0,1)) %>%
#                     left_join(., yDayAll) %>%
#                     mutate()

pred <- predict(mod6_YDCcr_Cr_CSr_Tr_Fr_TFr, newdata = predGrid)#, type = 'lpmatrix')
#predBKT2 <- predict(mod1, newdata = predGridBKT2)#, type = 'lpmatrix')

predGrid2 <- bind_cols(predGrid, as.data.frame(pred)) %>% mutate(pred = pred)
#predGrid2BKT2 <- bind_cols(predGridBKT2, as.data.frame(predBKT2)) %>% mutate(pred = predBKT2, species = 'bkt')

ggplot(predGrid2 %>% filter(salmon01 == 0),

       aes(ydayCumul, pred, group = cohort, color = cohort)) +
  #geom_point() + 
  geom_vline(xintercept = c(365,365*2), color = 'grey60', linetype = 2) +
  geom_line(size = 1.1) +
  labs(y = 'Predicted body size (mm)',
       x = 'Days since age-0 Jan 1',
       color = 'Cohort') +
  scale_color_viridis_c(breaks = c(2000,2004,2008,2012), labels = c(2000,2004,2008,2012)) +
  theme_publication() +
  theme(legend.position = "right", 
        legend.direction = "vertical",
        legend.title = element_text(face = "plain")) +
  guides(color = guide_colorbar(barheight = 10,
                                barwidth = 1.5,
                                ticks = FALSE,
                                nbin = 100)) +
#  guide_legend(title = 'Cohort') +
  facet_wrap(~riverFactor)#, scales = 'free_y')

@@

#using salmon01 * river because we need diff effects per river - even though AIC is worse for this model
modBKT <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor), 
             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")

@@

plot(modBKT, pages = 8, scale = 0)

predGridBKT <- expand.grid(riverFactor = unique(modBKT$model$riverFactor), 
                           cohortFactor = unique(modBKT$model$cohort),
                           ydayCumul = seq(250,1000, by = 10),
                           salmon01 = c(0,1),
                           cohortStrengthScaled = 0, #seq(-1.5,1.5, by = 0.5),
                           meanTemperatureScaledBySeason = 0, #seq(-1.5,1.5, by = 0.5),
                           meanFlowScaledBySeason = 0) %>%
                    left_join(., yDayAll) %>%
                    mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
                           seasonRiverFactor = paste(season, riverFactor, sep = '_'),
                           cohort = cohortFactor,
                           #river = riverFactor,
                           meanTemperatureFlowScaledBySeason = meanTemperatureScaledBySeason * meanFlowScaledBySeason)

@@

modBNT <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor), 
             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bnt'), method = "REML")

@@

predGridBNT <- expand.grid(riverFactor = unique(modBNT$model$riverFactor), 
                           cohortFactor = unique(modBNT$model$cohort),
                           ydayCumul = seq(250,1000, by = 10),
                           salmon01 = c(0,1),
                           cohortStrengthScaled = 0, #seq(-1.5,1.5, by = 0.5),
                           meanTemperatureScaledBySeason = 0, #seq(-1.5,1.5, by = 0.5),
                           meanFlowScaledBySeason = 0) %>%
                    left_join(., yDayAll) %>%
                    mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
                           seasonRiverFactor = paste(season, riverFactor, sep = '_'),
                           cohort = cohortFactor,
                           #river = riverFactor,
                           meanTemperatureFlowScaledBySeason = meanTemperatureScaledBySeason * meanFlowScaledBySeason)

@@

predGrid2Both$riverGG <- factor(predGrid2Both$riverFactor,
                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
                         labels = c("West Brook","Open Large","Open Small","Isolated Small"))
predGrid2Both$speciesGG <- factor(predGrid2Both$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"))

@@

# cohorts by river/species
ggplot(predGrid2Both %>% 
         filter(salmon01 ==0
#         (cohortStrengthScaledBKT == 0 & salmon01 == 0) | (cohortStrengthScaledBKT == 0 & cohortStrengthScaledBNT == 0 & cohortStrengthScaledATS == 0 & metaPop01 == 1 & salmon01 ==0)
           ),
#           cohortStrengthScaledBKT == 0, cohortStrengthScaledBNT == 0, cohortStrengthScaledATS == 0, metaPop01 == 1, salmon01 == 0),
       aes(ydayCumul, pred, group = cohort, color = cohort)) +
  #geom_point() + 
  geom_vline(xintercept = c(365,365*2), color = 'grey60', linetype = 2) +
  geom_line(size = 1.1) +
  labs(y = 'Predicted body size (mm)',
       x = 'Days since age-0 Jan 1',
       color = 'Cohort') +
  scale_color_viridis_c(breaks = c(2000,2004,2008,2012), labels = c(2000,2004,2008,2012)) +
  theme_publication() +
  theme(legend.position = "right", 
        legend.direction = "vertical",
        legend.title = element_text(face = "plain")) +
  guides(color = guide_colorbar(barheight = 10,
                                barwidth = 1.5,
                                ticks = FALSE,
                                nbin = 100)) +
#  guide_legend(title = 'Cohort') +
  facet_grid(speciesGG ~ riverGG, scales = 'free_y')

@@

```{r models with season-don't run. gam structure given best model from above}

AIC(mod0_YDCcr, mod0_YDCc_r, mod0_YDCcr_Cr_CSr_Tr_Fr_TFr, mod0_YDCcr_Cr_CSr_Ts_Fs_TFs,mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_C_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_CS_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_CSr,mod0_YDCcr_C_CS_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_Tsr_Fsr_TFsr,mod0_YDCcr_CSr_Tsr_Fsr_TFsr, mod0_YDCcr_Tsr_Fsr_TFsr, mod6_YDCcr_Cr_CSr_Tsr_Fsr_TFsr) %>% rownames_to_column() %>% arrange(AIC)

#AIC(mod6_YDCcr_Cr_CSr_Tsr_Fsr_TFsr)
#[1] 401367.4

#> AIC(mod6_YDCcr_Cr_CSr_Tr_Fr_TFr)
#[1] 402517.3

#########
# models with seasonal effects give a drop in seqson 1 sizes try models without seasonal effects
# may make sense because season is implicit in yDayCumul

# model without season in fixed
# this one gives a bump up in sizes for season 4
system.time(
mod6_YDCcr_Cr_CSr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
               data = dataIn, 
               method = "REML")
)

# predictions from this models are smooth!
# use this as the base model
system.time(
mod6_YDCcr_Cr_CSr_Tr_Fr_TFr <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor), 
               data = dataIn, 
               method = "REML")
)

summary(mod6_YDCcr_Cr_CSr_Tr_Fr_TFr) 
plot(mod6_YDCcr_Cr_CSr_Tr_Fr_TFr, pages = 8, shade = TRUE, scale = 0)
